---
title: "Assignment on linear models"
output: html_notebook
---

Anadarko Petroleum Corp. (APC) has been chosen to complete the assignment.

The data is obtained from Google Finance for a period since 2009/01/01, each observation is a single day.

**All statistical assumptions and conclusions are made on the 5% significance level.**

The closing price is obtained the following way:
```{r message=FALSE}
library(quantmod)
apc = Cl(getSymbols('RDY', src='google', auto.assign=FALSE, from='2009-01-01'))
```

# Stationary AR models
## Task (a)

> Compute and plot the log price xt and the log return rt. Comment on the two plots
(how volatile the data are, volatility clustering, outliers etc).

```{r}
logapc = log(apc)
logrtapc = diff(logapc)

plot(logapc)
plot(logrtapc)
```

## Task (b)

> Compute and plot the first 12 lags of ACF of xt. Comment on the plot. Based on the
ACF, is there a unit root in xt dataset? Why?

```{r}
acf(as.numeric(na.omit(logapc)))
```

## Task (c)

> Consider the time series for rt. Perform the Ljung-Box test for m = 12. Draw a
conclusion and justify it with the statistical language, i.e., in terms of the critical
region or p-value.

```{r}
Box.test(as.numeric(logrtapc), lag=12, type='Ljung-Box')
```

## Task (d)

> Use the command ar(rt,method='mle',order.max=20) to specify the order of an AR
model for rt. Use the PACF and AIC criteria (ar() and pacf() commands). Compare
both approaches.

```{r}
ar(as.numeric(na.omit(logrtapc)),method='mle',order.max=20)
```

```{r}
ar2$aic
```

```{r}
pacf(as.numeric(na.omit(logrtapc)))
```

## Task (e)

> Build an AR model for rt. Plot the time series of the residuals, ACF and p-values of
the Ljung-Box test (command tsdiag()). Perform the Ljung-Box test of the residuals
by hand adjusting the degrees of freedom for the number of the model parameters (see
[2], p.66). Is the model adequate? Why? Refine the model by eliminating all estimates
with t-ratio less than 1.645 and check the new model as described above. Is the new
model adequate? Why? Write down the final model.

```{r}
ar2 = arima(as.numeric(na.omit(logrtapc)), c(2,0,0))
ar2

ar2 = arima(as.numeric(na.omit(logrtapc)), c(2,0,0), fixed=c(0, NA, NA))
ar2
```

```{r fig.width=12}
tsdiag(ar2)
```

