---
title: "Assignment on linear models"
output: html_notebook
---

## Ксюша, шоколад должен быть белый с ванилью))

Anadarko Petroleum Corp. (APC) has been chosen to complete the assignment.

The data is obtained from Google Finance for a period since 2009/01/01, each observation is a single day.

**All statistical assumptions and conclusions are made on the 5% significance level.**

The closing price is obtained the following way:
```{r message=FALSE}
library(quantmod)
apc = Cl(getSymbols('RDY', src='google', auto.assign=FALSE, from='2009-01-01'))
```

# Stationary AR models
## Task (a)

> Compute and plot the log price xt and the log return rt. Comment on the two plots
(how volatile the data are, volatility clustering, outliers etc).

```{r}
logapc = log(apc)
logrtapc = diff(logapc)

plot(logapc)
plot(logrtapc)
```

## Task (b)

> Compute and plot the first 12 lags of ACF of xt. Comment on the plot. Based on the
ACF, is there a unit root in xt dataset? Why?

```{r}
acf(as.numeric(na.omit(logapc)))
```

## Task (c)

> Consider the time series for rt. Perform the Ljung-Box test for m = 12. Draw a
conclusion and justify it with the statistical language, i.e., in terms of the critical
region or p-value.

```{r}
Box.test(as.numeric(logrtapc), lag=12, type='Ljung-Box')
```

## Task (d)

> Use the command ar(rt,method='mle',order.max=20) to specify the order of an AR
model for rt. Use the PACF and AIC criteria (ar() and pacf() commands). Compare
both approaches.

```{r}
ar(as.numeric(na.omit(logrtapc)),method='mle',order.max=20)
```

```{r}
ar2$aic
```

```{r}
pacf(as.numeric(na.omit(logrtapc)))
```

## Task (e)

> Build an AR model for rt. Plot the time series of the residuals, ACF and p-values of
the Ljung-Box test (command tsdiag()). Perform the Ljung-Box test of the residuals
by hand adjusting the degrees of freedom for the number of the model parameters (see
[2], p.66). Is the model adequate? Why? Refine the model by eliminating all estimates
with t-ratio less than 1.645 and check the new model as described above. Is the new
model adequate? Why? Write down the final model.

```{r}
ar2 = arima(as.numeric(na.omit(logrtapc)), c(2,0,0))
ar2

ar2 = arima(as.numeric(na.omit(logrtapc)), c(2,0,0), fixed=c(0, NA, NA))
ar2
```

```{r fig.width=12}
tsdiag(ar2)
```

## Task (f)

> Does the model imply existence of a cycle? Why? If the cycles are present, compute
the average length of these cycles.

```{r}
k = predict(ar2, 5)
k
```

# MA model

## Task (a)

> Choose the order of such model. Support your choice with the ACF plot.

```{r}
acf(as.numeric(na.omit(logrtapc)))
```

## Task (b)

> Build the model. Refne it by removing coefficients estimates with t-ratio less than
1.645. Write down the ftted model.

```{r fig.width=12}
ma2 = arima(as.numeric(na.omit(logrtapc)), order=c(0,0,2))
ma2

ma2 = arima(as.numeric(na.omit(logrtapc)), order=c(0,0,2), fixed=c(0, NA, NA))
ma2

tsdiag(ma2)
```

## Task (c)

> Compute the Ljung-Box statistic of the residuals of the fitted MA model. Is there serial
correlation in the residuals? Why?

```{r fig.width=12}
Box.test(ma2$residuals, lag=12, type='Ljung-Box')
```

## Task (d)

> Consider the in-sample fits of the AR model of Problem 1 and the MA model. Which
model is preferred? Why?

<!-- Something about the AIC goes here, because otherwise they deal with the data quite well -->

## Task (e)

> Use backtest at some forecast origin with horizon h = 1 to compare the two models.
Indicate clearly the parameters of such backtesting (the estimation and forecasting
subsamples, forecast origin and so on). Which model is preferred? Why?

<!-- I guess I need to check the lecture slides -->

# ARIMA model

> 